---
title: "Reroducible Research - Course Project 1"
author: "Rassul Kalfane"
date: "29/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the dataLoading and preprocessing the data

First, we load and see some summary statistics of the data: 


```{r cars}
#Code for reading in the dataset and/or processing the data

unzip("repdata_data_activity.zip")
data <- read.csv("activity.csv")
summary(data)
names(data)
head(data)
pairs(data)
```

## What is mean total number of steps taken per day?

### 1- Calculate the total number of steps taken per day

```{r cars2}
stepsDay<-aggregate(steps ~ date, data, sum, na.rm=TRUE)
```


### 2- Make a histogram of the total number of steps taken each day

```{r cars3}
hist(stepsDay$steps)
```

### 3- Calculate and report the mean and median of the total number of steps taken per day

```{r cars4}
#Mean and median number of steps taken each day

meanStepsDay<-mean(stepsDay$steps)
meanStepsDay

medianStepsDay<-median(stepsDay$steps)
medianStepsDay
```


## What is the average daily activity pattern?

### 1- Make a time series plot 

```{r cars5}
stepsInterval<-aggregate(steps~interval, data=data, mean, na.rm=TRUE)
plot(stepsInterval)
```

### 2- Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r cars6}
maxInterval<-stepsInterval[which.max(stepsInterval$steps),]$interval
maxInterval
```

## Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA\color{red}{\verb|NA|}NA). The presence of missing days may introduce bias into some calculations or summaries of the data.  

### 1- Calculate and report the total number of missing values in the dataset 

numberNA<-sum(is.na(data$steps))
numberNA

### 2- Devise a strategy for filling in all of the missing values in the dataset.  

The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r cars7}
intervalMean<-function(interval){
  stepsInterval[stepsInterval$interval==interval,]$steps
}
```

### 3- Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r cars8}
dataWithoutNA<-data
for(i in 1:nrow(dataWithoutNA)){
  if(is.na(dataWithoutNA[i,]$steps)){
    dataWithoutNA[i,]$steps <- intervalMean(dataWithoutNA[i,]$interval)
  }
}
```


### 4- Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r cars9}
stepsDayWithoutNA <- aggregate(steps ~ date, data=dataWithoutNA, sum)
hist(stepsDayWithoutNA$steps)
```


## Are there differences in activity patterns between weekdays and weekends?

### 1- Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r cars10}
#Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

dataWithoutNA$date <- as.Date(strptime(dataWithoutNA$date, format="%Y-%m-%d"))
dataWithoutNA$day <- weekdays(dataWithoutNA$date)
for (i in 1:nrow(dataWithoutNA)) {
  if (dataWithoutNA[i,]$day %in% c("Saturday","Sunday")) {
    dataWithoutNA[i,]$day<-"weekend"
  }
  else{
    dataWithoutNA[i,]$day<-"weekday"
  }
}
stepsByDay <- aggregate(dataWithoutNA$steps ~ dataWithoutNA$interval + dataWithoutNA$day, dataWithoutNA, mean)
```

### 2- Make a panel plot containing a time series plot

```{r cars11}
#Make a panel plot containing a time series plot (i.e. type = "l"\color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

names(stepsByDay) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps ~ interval | day, stepsByDay, layout = c(1, 2), 
       xlab = "Interval", ylab = "Number of steps")
```